---
title: "Assignment 4"
author: "Balaji Ravi Kumar"
date: "19/03/2022"
output:
  html_document: default
  pdf_document: default
---

```{r}
#loading libraries required for the assignment
library(tidyverse)
library(factoextra)
library(cluster)
library(ggplot2)
library(gridExtra)
library(readr)
#importing dataset into R
pharmaceuticals<-read.csv("Pharmaceuticals.csv")
summary(pharmaceuticals)


```

```{r}
# a. Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on. 

# removeing any missing value(NA) that might be present in the Pharmaceutical data and selecting numerical data
pharmaceuticals1 <- na.omit(pharmaceuticals)
pharmaceuticals1
row.names(pharmaceuticals1)<- pharmaceuticals1[,1]
pharmadata<- pharmaceuticals1[, 3:11]
head(pharmadata)
```


```{r}
# Scaling the Pharmadata using the scale function
pharmascale <- scale(pharmadata)
head(pharmascale)
```

```{r}
#Determining the number of clusters by cluster analysis using Elbow and Silhouette Method

fviz_nbclust(pharmascale, kmeans, method = "wss") + labs(subtitle = "Elbow Method")



fviz_nbclust(pharmascale, kmeans, method = "silhouette")+ labs(subtitle = "Silhouette Method")



```


```{r}
#From the above plots, it is clear that the no of clusters are 5 and they are enough to
#show the variations that are present in the data

set.seed(64060)
#Applying kmean
kmeans5<- kmeans(pharmascale,centers=5,nstart = 25)

#Visualizing the output

#centroids

kmeans5$centers
fviz_cluster(kmeans5,data = pharmascale) # to Visualize the clusters
kmeans5
distance<- dist(pharmascale, method = "euclidean")
fviz_dist(distance)
```

```{r}
#K-Means Cluster Analysis Fitting the data with 5 clusters

clusterfit<-kmeans(pharmascale,5)


#Finding the mean value of all quantitative variables for each cluster

aggregate(pharmascale,by=list(clusterfit$cluster),FUN=mean)

pharmadata1<-data.frame(pharmascale,clusterfit$cluster)
pharmadata1
View(pharmadata1)

## Final analysis and Extracting results using 5 clusters and Visualize the results


clusplot(pharmascale,clusterfit$cluster,color = TRUE,shade = TRUE,labels = 2,lines = 0)
```



#b Interpret the clusters with respect to the numerical variables used in forming the clusters
```{r}
Pharma_Cluster <- pharmaceuticals[,c(12,13,14)]%>% mutate(clusters = clusterfit$cluster)%>% arrange(clusters, ascending = TRUE)
Pharma_Cluster


#Cluster1 - JNJ,MRK,GSK,PFE 

#Cluster2 - AHM,AVE,WPI 

#Cluster3 - IVX,MRX,ELN,CHTT

#Cluster4 - AGN,PHA,BAY 

#Cluster5 - ABT,NVS,AZN,LLY,BMY,WYE,SGP

```


# c)Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)
#There is a pattern in the clusters with respect to the recommendation variable.
#Cluster 1 , which has the highest Market capitalization, highest ROA, highest ROE and high Asset_Turnover has no median sales recommendations.Cluster 1 has equal hold and Moderate buy recommendation.
#Cluster 2, with the highest Revenue growth has Lowest P/E Ratio and lowest Asset turn over has buy and sell recommendation.
#Cluster 3, which has highest revenue growth with lowest Market capitalization,lowest ROE, ROA, Leverage, revenue growth and net profit margin has mostly buy recommendation
#Cluster 4, with highest P/E ration and lowest leverage and asset turnover has hold recommendation
#Cluster 5, which has the highest net profit margin with lowest leverage and beta has mostly hold recommendation.

# d)Provide an appropriate name for each cluster using any or all of the variables in the dataset.

Cluster-1 - Moderate buy/Hold cluster.

Cluster-2 - Buy/Sell cluster.

Cluster-3 - Buy Cluster.

Cluster-4 - Hold cluster

Cluster-5 - High Hold cluster


